Figures to insert in text

### 2018-04-09 Chapters/Chapter02.tex
comparison

Consider the implementations of SDM present --- as of this writing --- on its wikipedia entry:

\begin{enumerate}
    \item The original hardware implementation developed by NASA [4]

    \item LISP implementation for the Connection Machine [60]

    \item APL implementation [59]

    \item FPGA implementation \citep{silva_reconfigurable_2004}

    \item C Binary Vector Symbols (CBVS): includes SDM implementation in C as a part of vector symbolic architecture[54] developed by EISLAB at Luleå University of Technology: http://pendicular.net/cbvs.php[55]

    \item CommonSense ToolKit (CSTK) for realtime sensor data processing developed at the Lancaster University includes implementation of SDM in C++: http://cstk.sourceforge.net/[56]

    \item Learning Intelligent Distribution Agent (LIDA) developed by Stan Franklin's lab at the University of Memphis includes implementation of SDM in Java: http://ccrg.cs.memphis.edu/framework.html

@@

It is our belief that a tool such as a framework could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as Python, in an intuitive platform such as Juypyter notebooks. Neuroscientists interested in long-term memory storage should not have to worry about high-bandwidth vector parallel computation.  This new tool would provide a ready to use system in which experiments could be executed almost as soon as designed --- and it might accelerate research \citep{shen2014interactive}.

Our motivation was our own effort to run our models. As there is no reference implementation, we had to develop our own and run several simulations to ensure that our implementation was correct and bug-free. Thus, we had to deviate from our primary goal --- which was to test our hypothesis and explore the `ideas space' --- and to focus on the implementation itself. Furthermore, new members of our research group had to go through different source code developed by former members, in different languages.

The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends and operations, (iii) has been validated against the mathematical model, (iv) is cross-platform, and (v) is easily extensible to test new research ideas.

Extensions of SDM have been used in many applications. For example, \citet{Snaider2011} extended SDM to store sequences of vectors and trees efficiently.  \citet{Rajesh1998} used a modified SDM in an autonomous robot. \citet{Meng2009} modified SDM to clean patterns from noisy inputs. \citet{fan1997genetic} extended SDM with genetic algorithms. \citet{chada2016you} extended SDM creating the Rotational Sparse Distributed Memory (RSDM), which is used to modeling network motifs, dynamic flexibility, and hierarchical organization, all results from neuroscience literature.

>>>>

\section{The wasted effort of duplicated work}

Our motivation is our own effort to run our models. As there is no reference implementation, we had to develop our own and run several simulations to ensure that our implementation was correct and bug-free. Thus, we had to deviate from our primary goal --- which was to test our hypothesis and explore the `ideas space' --- and to focus on the implementation itself. Furthermore, new members of our research group had to go through different source code developed by former members, in different languages.  Consider the implementations of SDM present --- as of this writing --- on its wikipedia entry\citep{noauthor_sparse_2018}:

\begin{enumerate}
    \item The original hardware implementation developed by NASA [4];

    \item LISP implementation for the Connection Machine [60];

    \item APL implementation [59];

    \item FPGA implementation \citep{silva_reconfigurable_2004};

    \item C Binary Vector Symbols (CBVS): includes SDM implementation in C as a part of vector symbolic architecture[54] developed by EISLAB at Luleå University of Technology: http://pendicular.net/cbvs.php[55];

    \item CommonSense ToolKit (CSTK) for realtime sensor data processing developed at the Lancaster University includes implementation of SDM in C++: http://cstk.sourceforge.net/[56];

    \item Learning Intelligent Distribution Agent (LIDA) developed by Stan Franklin's lab at the University of Memphis includes implementation of SDM in Java: http://ccrg.cs.memphis.edu/framework.html;

@@

Then there is LIDA --- a whole cognitive architecture based on Hofstadter's Fluid Concepts, Kanerva's SDM, and other ideas.  It is developed in Java; which makes it difficult to connect to the lowest levels of hardware; to connect to GPUs or FPGAs, and to other languages --- at least in comparison to Python\footnote{Python is sometimes called a `glue language'. That is, in my opinion, not the best metaphor.  A glue connects two things leaving an inflexible structure.  Python is perhaps best described as the interstate highway system of Programming; if something is out there, there is a way to reach it with Python.  In the comparison with Java, for instance, take the $popcnt(xor(b_i,b_j))$ operation, executed billions of times in SDM. How easy is it to program that for a particular GPU or FPGA with Java?}.  It is not available on an open-source repository.  It has a non-standard license, strange to the open-source community, ``the LIDA Framework Software NonExclusive, Non-Commercial Use License''.  Potential contributors must sign an ``Agreement Regarding Contributory Code for the LIDA Framework Software''... \emph{`before Memphis can accept it'}.

The closest implementations to ours, in philosophy at least, is the one in `the common sense toolkit'. It is executed in C++, with a normal open-source license, and hosted on an open-source code repository.  It is, however, strikingly dissimilar to ours on the following aspects:

\begin{enumerate}
    \item SDM is but a part of the system; the description of the system reads that cstk is `A toolkit for processing and visualising sensor data in real time with support for use with embedded platforms.'

    \item The whole SDM code is composed of 143 \emph{lines} of C++ in the $cstk/cstk-devonly/sdm$ folder.
0
    \item There is no work on making the system parallel.

    \item There is strong coupling between location address and location data, which makes experimentation hard.

    \item There are no tests or examples to be found instantly.

    \item Finally, the last commit to this repository seems to have been made in 2005?

    \item There are no publishable / published scientific applications or experiments available to be reproduced.

    \item there is no tutorial, installation instructions, performance benchmarks, framework validation or SDM Documentation.

Note that all these criticisms apply to the implementations in both the `common sense toolkit' and the `C Binary Vector Symbols' REF.  While these implementations have around 150 lines of C++; at last count, the \emph{documentation of our implementation} had over 100 \emph{pages}. We aim a higher here, and, to us, SDM is at center stage.

There is obviously a demand for use of SDM; but each group has been tied to their own \emph{ad-hoc} needs, and there has not been the emergence of a community centered on a tool. It is our belief that a tool such as standard open-source framework could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as Python, in an intuitive platform such as Juypyter notebooks. Neuroscientists interested in long-term memory storage should not have to worry about high-bandwidth vector parallel computation.  This new tool would provide a ready to use system in which experiments could be executed almost as soon as designed --- and provide the needed replication of studies \citep{shen2014interactive}.

The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends\footnote{CPUs, GPUs, FPGAs} and operations, (iii) is fully validated against the mathematical model, (iv) is cross-platform\footnote{Unix, Linux, MacOs, Windows, Amazon Web Services, etc.}, and (v) is easily extensible to test new research ideas --- and to let others replicate the results.

Another issue is \emph{extensibility}: Extensions of SDM have been used in many applications. For example, \citet{Snaider2011} extended SDM to store sequences of vectors and trees efficiently.  \citet{Rajesh1998} used a modified SDM in an autonomous robot. \citet{Meng2009} modified SDM to clean patterns from noisy inputs. \citet{fan1997genetic} extended SDM with genetic algorithms. \citet{chada2016you} extended SDM creating the Rotational Sparse Distributed Memory (RSDM), which is used to modeling network motifs, dynamic flexibility, and hierarchical organization, all results from neuroscience literature.
